Project Thesis/Report

### 2021-12-31 Gandhi_Project_Report/draft1.tex
Section 4 - just before memloss diagram. Need to split Th.4

% \subsection{Taken Embedding Theorem}

@@

\begin{Example} \label{Ex_exp}
 Consider the driven system $g(u_n,x_n)=x^n$ for $X=[0,1]$, $U=\mathbb{R}$. The system has an uncountable number of solutions, as there exists a solution for every $x\in{X}$ which also passes through the point $x$ and $lim_{n\to\infty}x_n=0$, $lim_{n\to{-}\infty}x_n=1$. 
 The proof is deferred to immediately after the next paragraph.

@@

\begin{Definition}
  [\bf Reachable Set] \label{Dfn_ReachableSet}
The reachable set of a driven system $g$ is exactly the union of all the elements of all the solutions, i.e., 
\[X_U :=\Big \{x \in X:  x = x_k \mbox{ where  $\{x_n\}$  is a solution for some  $\bar{u}$} \Big \}.\]
The set of all reachable states at a specific time $n$ for input $\overline{u}$ is denoted by $X_n(\overline{u})$
\end{Definition}

@@

\begin{Definition}
  [\bf Topological Contraction] \label{Dfn_TopContr}
  A function $g:U\times{X}\to{X}$ is a topological contraction if for all $n\in\mathbb{Z}$ and all $\overline{u}\subseteq{U}$, $X_n(\overline{u})$ is a singleton subset of X.  
\end{Definition}

\begin{proof}
  [\bf Proof of Example \ref{Ex_exp}] 
  To see this, we show that for every input $\overline{u}=\{u_n\}_{n\in\mathbb{Z}}$, $g(u_n,\cdot)$ is a contraction map on $X$ for every $k\in\mathbb{Z}$. 
  Indeed $g(u_n,x_n)=x^n$ depends only on the state at time $n$ and so for each input $\overline{u}$, $X_n(\overline{u})=\{x^n\}$ is a singleton subset.
  We then easily conclude that $\{x_n\}\equiv0$ is the only solution. 
\end{proof}

@@

%%%%%%%%%%%%%%%%%%%%%ISSUE HERE - fix before sending
\begin{Theorem} \label{Thm_CompactExistence}
  If $X$ is compact then for each input $\overline{u}$, there exists at least one solution to the driven system $g(\ldotp, x)$
\end{Theorem}
\begin{proof}
  Consider an input $\overline{u}={\{u_n\}}_{n\in\mathbb{Z}}\subseteq{U}$ and driven system $g:U\times{X}\to{X}$  generating a sequence ${\{x_n\}}_{n\in\mathbb{Z}}$ in the compact space $X$. 
  Since $X$ is a metric space, it follows immediately from a well-known result in Analysis (\textbf{Cite}) that $\{x_n\}$ has a convergent subsequence which then is a solution. 
\end{proof}

@@

\begin{Definition}
  [\bf Unique Solution Property)] \label{Dfn_usp}
  A driven system $g$ is said to have the Unique Solution Property (USP) if for each input $\overline{u}$ there exists exactly one solution. 
  Alternatively we may formulate the USP as follows: $g$ has the Unique Solution Property if there exists a well-defined map $\Psi:{U}\to{X}$ with $\Psi({\overline{u}})$ denoting the unique solution.
\end{Definition}

@@

  \begin{Definition} 
    [\bf Universal Semi-Conjugacy)] \label{Def_UnivSemiConj}
    Given a driven system $g$, we  call a continuous and surjective map $h : \overleftarrow{U} \to X_U$ a universal semi-conjugacy if  diagram \ref{Scomm_h} commutes for all $v \in U$.
  \end{Definition}

@@

Re-sketching the graph \ref{Scomm_h} above by fixing the input $v$ in $g$ and replacing $X_U$ by its left-infinite sequence space $\overleftarrow{U}$, we obtain the graph below. In this case, the function $H:\overleftarrow{U}\to\overleftarrow{X}_U$, a map that is both continuous and surjective, is called a \emph{causal mapping}.

@@

  Note that $\tilde{g}_v$ maps $(\ldots, u_{-2}, u_{-1})$ to $(\ldots, u_{-2}, u_{-1}, g(v, u_{-1}))$.
  The driven system $g$ can induce an embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ as follows:

\begin{Definition}
  [\bf Causal Embedding]\label{Def_CausEmbed}
  If $H:\overleftarrow{U}\to\overleftarrow{X}_U$ is also injective (in addition to being surjective and hence bijective), it becomes the embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ induced by the driven system $g$ and we refer to $H$ as a \emph{causal embedding}. 
\end{Definition}

\begin{Theorem}
  For a compactly driven system, a causal mapping $H$ exists if and only if $g$ has the USP. 
\end{Theorem}
\begin{proof}
  \emph{See proof from p.15 of \cite{manjunath2021universal}}
\end{proof}

Note that even when $h$ does exist, we are not guaranteed its injectivity. Considering again example \ref{ex_halfux}, we conclude that even if $h$ were to exist, it could not be injective as $X_U=\{0\}$.

@@

  Before formalising this, we make mention of the concept of the \emph{inverse-limit system} of a system $(U,T)$. In broad sweeps, the inverse-limit system of a dynamical system is a self-map on a subset of an infinite-dimensional space \textbf{(expand what exactly this is?)} where each point in the inverse-limit space corresponds to a backward orbit of the map $T$. In literature, the inverse-limit space is denoted by $\hat{U}_T\subseteq\overleftarrow{U}$. \textbf{cite from supp.koopman}. 
  The map $T$, then, induces a self-map $\hat{T}:\widehat{T}: (\ldots,u_{-2},u_{-1}) \mapsto  (\ldots,u_{-2},u_{-1},T(u_{-1}))$ on $\widehat{U}_T$

\begin{Definition}
  A driven system $g$ is said to causally embed the dynamical system $(U,T)$ if 
  \vspace{-8mm}
\begin{enumerate}[noitemsep, label=\roman*.]
  \item The diagram below commutes (i.e. $g$ is a universal semi-conjugacy),
  \item $H_2(\overline{u}):=(h(r\overline{u}, h(\overline{u})))$ embeds the inverse-limit space $(\hat{U}, \hat{T})$ of $(U,T)$ in $X\times{X}$.
\end{enumerate}
\end{Definition}

\begin{Theorem}
 If $g(\ldotp,x)$ is invertible and has the USP then $H$ is a causal embedding. 
\end{Theorem}

When g has the USP, the diagram below illustrates the operation of the the mappings h and H the mapping $h:\overleftarrow{U}\to{X_U}$ is an observable as discussed (above).

@@

\begin{Theorem}
  The following statements are equivalent:
  \vspace{-8mm}
  \begin{enumerate}[noitemsep, label=\roman*.]
    \item $g$ has the USP 
    \item $g$ is topological contraction 
    \item $g$ has the Unique Attraction Property (UAP). 
  \end{enumerate}
\end{Theorem}
\begin{proof}
  (we’ll show this)  [11] eqn.6 
\end{proof}

We have not as of yet treated the UAP, and establish the necessary result here (without proof - left to literature. Or later) so as not to break up the completeness of theorem

It is important to note here that one must be careful to avoid a choice of g which adds additional complexity to the obtained solution.  
When the causal embedding exists for the driven system $g$, one may map an arbitrary input $u$ onto the solution space $X$ without additional distortion or information-loss.  
Establishing an embedding removes this problem by guaranteeing that, since the systems are conjugate (semi-conjugate), $g$ does not add any (only “some”) complexity to the system.  
It is, however, a balancing act as one also does not want to choose a function g that quenches the temporal structure in $u$ by contracting to such a degree that one loses the ability to recover information from the original system completely. To obtain this we want the reachable set of a driven system to be large enough to relate to the input.

In the example above \ref{ex_halfux}, we cannot relate the input's temporal variation and the reachable set because $X_U$ consists of a single element.

The reachable set $Y_T$ of a driven system must therefore be such that we can embed the inverse-limit space of $U_T$ can be embedded in some finite self-product of the reachable set of $T$. To this end, consider the notion of State-Input (SI) Invertibility.

\begin{Definition}
  A map $g$ is said to be SI-Invertible if $g$ is invertible for all $x\in{X}$. Alternatively, if, given $x_n$ and $x_{n-1}$, $u_n$ can be uniquely determined from $x_n=g(u_n,x_{n-1})$, then $g$ is said to be SI-invertible.
\end{Definition}

We make special note of a specific driven system. The function $$g(u,x)=(1-a)x+atanh(Au+\alpha{a}Bx)$$  is both SI-invertible and possesses the USP. This function $g$ is used in our implementation and is more completely discussed in chapter 5 \ref{sect5}
The proof of these two facts is rather involved and the proof, instead of being replicated here, may be found in (cite relevant article)

\textbf{Expand: UAP + Embedding -> perturbation isn’t an issue + we can drop left-infinite sequence.  }

Despite the ease that one may work with left-infinite sequences in the realm of theory, it is impossible to obtain such a sequence in any real-life application.  One does not in practice, however, need the entire left-infinite history of the input due to the Uniform Attraction Property (UAP). The next definition is stated in an informal manner as the more rigourous definition makes use of processes, a concept which would take some amount of paper to establish and explain - something which would detract from the overarching thrust of this paper/project/thesis. \textbf{is it VERY obsious to the reader at this point what the general 'thrust' of the paper IS?} \textbf{Cite articles}

\begin{Definition}
  A driven system $g$ has the Uniform Attraction Property (UAP) if, regardless of starting position, all trajectories converge to a single trajectory as time flows forward. 
  This trajectory is also the unique solution sequence $x$ to the input sequence $u$ as mentioned above \ref{here_UAP_firstestablished}.
\end{Definition}

Incredibly, one may then initialise a driven system with a completely arbitrary initial value $y_m\in{X}$ where $m\in\mathbb{Z}$ and the UAP then guarantees that the sequence $\{y_{m+1}, y_{m+2},\ldots\}$ satisfying $y_n=g(u_n, y_{n-1})$  with $k\geq{m}$  will uniformly approach the elements $\{x_n\}$ of the actual solution. (see [41, Theorem 1] or [11, Eqn. (18)])

@@

\begin{figure}[h]
  \includegraphics[scale=0.4]{memloss_conttime.png}
  \centering
  \label{fig:memoryLosscont_time}
%\caption{Approaching the true solution if g has the UAP}
\end{figure}

This also solves the problem of perturbations or noise introduced by the observable or measurement function. \textbf{(rephrase)}

>>>>

\subsection{Taken Embedding Theorem} \label{subs_TakensEmbedding}

@@

\begin{Example}\label{Ex_exp}
 Consider the driven system $g(u_n,x_n)=x^n$ for $X=[0,1]$, $U=\mathbb{R}$. The system has an uncountable number of solutions, as there exists a solution for every $x\in{X}$ which also passes through the point $x$ and $lim_{n\to\infty}x_n=0$, $lim_{n\to{-}\infty}x_n=1$. 
 The proof is deferred to immediately after the next paragraph.

@@

\begin{Definition}
  [\bf Reachable Set]\label{Dfn_ReachableSet}
The reachable set of a driven system $g$ is exactly the union of all the elements of all the solutions, i.e., 
\[X_U :=\Big \{x \in X:  x = x_k \mbox{ where  $\{x_n\}$  is a solution for some  $\bar{u}$} \Big \}.\]
The set of all reachable states at a specific time $n$ for input $\overline{u}$ is denoted by $X_n(\overline{u})$
\end{Definition}

@@

\begin{Definition}
  [\bf Topological Contraction]\label{Dfn_TopContr}
  A function $g:U\times{X}\to{X}$ is a topological contraction if for all $n\in\mathbb{Z}$ and all $\overline{u}\subseteq{U}$, $X_n(\overline{u})$ is a singleton subset of X.  
\end{Definition}

\begin{proof}
  [\bf Proof of Example\ref{Ex_exp}] 
  To see this, we show that for every input $\overline{u}=\{u_n\}_{n\in\mathbb{Z}}$, $g(u_n,\cdot)$ is a contraction map on $X$ for every $k\in\mathbb{Z}$. 
  Indeed $g(u_n,x_n)=x^n$ depends only on the state at time $n$ and so for each input $\overline{u}$, $X_n(\overline{u})=\{x^n\}$ is a singleton subset.
  We then easily conclude that $\{x_n\}\equiv0$ is the only solution. 
\end{proof}

@@

\begin{Theorem}\label{Thm_CompactExistence}
  If $X$ is compact then for each input $\overline{u}$, there exists at least one solution to the driven system $g(\ldotp, x)$
\end{Theorem}
\begin{proof}
  Consider an input $\overline{u}={\{u_n\}}_{n\in\mathbb{Z}}\subseteq{U}$ and driven system $g:U\times{X}\to{X}$  generating a sequence ${\{x_n\}}_{n\in\mathbb{Z}}$ in the compact space $X$. 
  Since $X$ is a metric space, it follows immediately from a well-known result in Analysis (\textbf{Cite}) that $\{x_n\}$ has a convergent subsequence which then is a solution. 
\end{proof}

@@

\subsection {Unique Solution Property}

\begin{Definition}
  [\bf Unique Solution Property] \label{Dfn_usp}
  A driven system $g$ is said to have the Unique Solution Property (USP) if for each input $\overline{u}$ there exists exactly one solution. 
  Alternatively we may formulate the USP as follows: $g$ has the Unique Solution Property if there exists a well-defined map $\Psi:{U}\to{X}$ with $\Psi({\overline{u}})$ denoting the unique solution.
\end{Definition}

@@

  \begin{Definition} 
    [\bf Universal Semi-Conjugacy]\label{Def_UnivSemiConj}
    Given a driven system $g$, we  call a continuous and surjective map $h : \overleftarrow{U} \to X_U$ a universal semi-conjugacy if  diagram \ref{Scomm_h} commutes for all $v \in U$.
  \end{Definition}

@@

Re-sketching the graph\ref{Scomm_h} above by fixing the input $v$ in $g$ and replacing $X_U$ by its left-infinite sequence space $\overleftarrow{U}$, we obtain the graph below. In this case, the function $H:\overleftarrow{U}\to\overleftarrow{X}_U$, a map that is both continuous and surjective, is called a \emph{causal mapping}.

\begin{Definition}
  [\bf Causal Mapping]\label{Def_CausMap}
  A continuous, surjective map $H:\overleftarrow{U}\to\overleftarrow{X}_U$ such that \[H\circ\tilde{g}_v=\sigma\circ{H}\] where $\tilde{g}_v$ maps $(\ldots, u_{-2}, u_{-1})$ to $(\ldots, u_{-2}, u_{-1}, g(v, u_{-1}))$.
\end{Definition}

@@

\begin{Theorem}
  For a compactly driven system, a causal mapping $H$ exists if and only if $g$ has the USP. 
\end{Theorem}
\begin{proof}
  \emph{See proof from p.15 of \cite{manjunath2021universal}}
  \textbf{Expand.}
\end{proof}

Note that even when $h$ does exist, we are not guaranteed its injectivity. Considering again example \ref{ex_halfux}, we see that even if $h$ were to exist, it could not be injective as $X_U=\{0\}$.

@@

  Before formalising this, we make mention of the concept of the inverse-limit system of a dynamical system $(U,T)$. 
  In broad sweeps, the inverse-limit system of a dynamical system is a self-map on a subset of an infinite-dimensional space \textbf{(Expand)} where each point in the inverse-limit space corresponds to a backward orbit of the map $T$. In literature, the inverse-limit space is denoted by $\hat{U}_T\subseteq\overleftarrow{U}$ \cite{manjunath2021universal}. 
  The map $T$, then, induces a self-map $\hat{T}:\widehat{T}: (\ldots,u_{-2},u_{-1}) \mapsto  (\ldots,u_{-2},u_{-1},T(u_{-1}))$ on $\widehat{U}_T$

\begin{Definition}
  [\bf Causal Embedding]\label{Dfn_CausalEmbed}
  A driven system $g$ is said to causally embed the dynamical system $(U,T)$ if 
  \vspace{-8mm}
\begin{enumerate}[noitemsep, label=\roman*.]
  \item The diagram in \ref{SCausal_H} commutes (i.e. $g$ is a universal semi-conjugacy),
  \item $H_2(\overline{u}):=(h(r\overline{u}, h(\overline{u})))$ embeds the inverse-limit space $(\hat{U}, \hat{T})$ of $(U,T)$ in $X\times{X}$.
\end{enumerate}
\end{Definition}

The driven system $g$ can induce an embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ as follows: If $H:\overleftarrow{U}\to\overleftarrow{X}_U$ is also injective (in addition to being surjective, and hence bijective), it becomes the embedding of $\overleftarrow{U}$ in $\overleftarrow{X}$ induced by the driven system $g$ and we refer to $H$ as a \emph{causal embedding}.

\begin{Theorem}
 If $g(\ldotp,x)$ is invertible and has the USP then $H$ is a causal embedding. 
\end{Theorem}
\begin{proof}
  \emph{See proof from \cite{manjunath2021universal}}  \textbf{Expand.}
\end{proof}

When g has the USP, the diagram below illustrates the operation of the the mappings $h$ and $H$. The mapping $h:\overleftarrow{U}\to{X_U}$ is an observable as discussed in an above section (\ref{TakensEmbedding}).

@@

The next result is of relative import but to fully comprehend its influence, some discussion still remains. Before stating it, we define a concept which will directly be shown to be equivalent to the USP.

\begin{Definition}
  [\bf Uniform Attraction Property]\label{Dfn_UAP}
  A driven system $g$ has the Uniform Attraction Property (UAP) if, regardless of starting position, all trajectories converge to a single trajectory as time flows forward. 
  This trajectory is also the unique solution sequence $x$ to the input sequence $u$ as mentioned above \ref{here_UAP_firstestablished}.
\end{Definition}

\begin{Theorem}
  The following statements are equivalent:
  \vspace{-8mm}
  \begin{enumerate}[noitemsep, label=\roman*.]
    \item $g$ has the USP 
    \item $g$ is a topological contraction 
    \item $g$ has the Unique Attraction Property (UAP). 
  \end{enumerate}
\end{Theorem}
\begin{proof}
  (we’ll show this)  [11] eqn.6 
\end{proof}

We have not as of yet treated the UAP properly, and establish the necessary result here (without proof) so as not to break up the completeness of theorem.

\textbf{The proof of the result for the UAP is left to literature. Why? Cite.}

\subsection {Choosing the driven system $g$}

One must be careful to avoid a choice of $g$ which would  add complexity to the obtained solution.

When a causal embedding $H$ exists for the driven system $g$, one can map an arbitrary input ${u}$ onto the solution space $X$ without additional distortion or information-loss.  \textbf{(Cite.)} 
When an embedding is established, the question of possible additional complexity in the solution is removed by guaranteeing that, since the systems are conjugate (semi-conjugate, (refer)), $g$ does not add any (some) complexity to the system.  
It is, however, a balancing act as it also undesirable to choose a function $g$ that quenches the temporal structure in $u$ by contracting to such a degree that the ability to recover information from the original system is lost completely. 
To obtain a suitably complex function $g$, it is desired that the the reachable set of a driven system  be large enough to relate to the input.  \textbf{(Expand here. It stops a bit abruptly.)}

In the example above (\ref{ex_halfux}), the input's temporal variation cannot be related to the reachable set as $X_U$ consists of a single element and so little, if not no, information is encoded.  Rather, the reachable set of a driven system must therefore be such that the inverse-limit space of $U_T$ can be embedded in some finite self-product of the reachable set of $T$ \textbf{(refer or cite)}. To this end, consider the notion of State-Input (SI) Invertibility.

\begin{Definition}
  [\bf SI-Invertibility]\label{Dfn_SIinv}
  A map $g$ is said to be SI-Invertible if $g$ is invertible for all $x\in{X}$. Alternatively it may be said that if, given $x_n$ and $x_{n-1}$, $u_n$ can be uniquely determined from $x_n=g(u_n,x_{n-1})$, then $g$ is said to be SI-invertible.
\end{Definition}

\textbf{Explain more here.}

We make special note of a specific driven system. The function

\begin{equation}  \label{eqn_driveng}
%    \[ 
g(u,x)=(1-a)x+a{\cdot}tanh(Au+\alpha{a}Bx)
% \]
\end{equation}

is both SI-invertible and possesses the USP. The proof of these two facts is rather involved and  hence the proof, instead of being replicated here, may be found in (cite relevant article) 
This specific driving function $g$ is used in our implementation and is more completely discussed in chapter \ref{sect5}

\textbf{Expand: UAP + Embedding $\rightarrow$  perturbation isn’t an issue + we can drop left-infinite sequence.  }

Despite the ease that one may work with a left-infinite history in the realm of theory, it is impossible to obtain such a sequence in any real-life application.  
One does not in practice, fortunately, need the entire left-infinite history of the input thanks to the Uniform Attraction Property (UAP). 
The next definition is stated in an informal manner as the more rigourous definition makes use of processes, a concept which would take some amount of paper to establish and prove - an exercise which would detract from the principal thrust of this paper/project/thesis.

\textbf{Is it VERY obvious to the reader at this point what the general 'thrust' of the paper IS?} \textbf{Cite articles}

Recall that the Unique Attraction Property guarantees that all trajectories \textit{(which ones?)} will converge to the same trajectory in forward time. 
Incredibly, this permits one to initialise a driven system $g$ with an altogether arbitrary initial value $y_m\in{X}$ where $m\in\mathbb{Z}$ and the UAP guarantees that the sequence $\{y_{m+1}, y_{m+2},\ldots\}$ satisfying the relation $y_n=g(u_n, y_{n-1})$  where $k\geq{m}$  will uniformly approach the elements $\{x_n\}$ of the actual solution. (see [41, Theorem 1] or [11, Eqn. (18)])

%Change the theorem above so that th.4 is split up into two parts.

\begin{figure}[ht]
  \includegraphics[scale=0.4]{memloss_conttime.png}
  \centering
  \label{fig:memoryLosscont_time}
%\caption{Approaching the true solution if g has the UAP}
\end{figure}

@@

This also solves the problem of perturbations or noise introduced by the observable or measurement function. \textbf{(rephrase, expand and explain)}

\subsection{The next step in Dynamics}
